An event-sourced and decentralized database with IPLD and materialized views

### 2019-09-23 draft.tex
update: cleanup, refs, better text
Signed-off-by: Carson Farmer <carson.farmer@gmail.com>

To make Threads as easy to adopt and use as possible, we've designed the internal event store to take advantage of ideas from several existing CQRS and ES systems, as well as ideas and designs from Flux\footnote{http://facebook.github.io/flux}/Redux\footnote{Flux/Redux build on concepts from CRQS and ES themselves, and Redux is arguably an implementation of the Flux application architecture (i.e., Flux is a pattern, Redux is a library).} \cite{reduxMotivation} and domain driven design [\dots] (DDD). In the following sections, we outline the internal structure of a local Threads ``client'' at a relatively high-level, and outline how it can be used to build intuitive developer-facing APIs to make adopting and using Threads ``the right choice'' for a wide range of developers.

@@

For application developers familiar with common application state management issues, tools such as Flux/Redux provide opinionated frameworks for making state mutations \emph{predictable} by imposing certain restrictions on how and when updates can happen. If you take a look at the architecture of a Flux (or Redux) application, you'll notice several core similaries to CQRS and ES systems (see Figure \dots). In both cases, the focus is on unidirectional data flows that build downstream views from atomic updates in the form of events (or actions).

We adopt a similar flow in Threads, such that views are updated by subscribing to specific events (or actions as in Flux/Redux), and applying a reducer function that is used to update the view state (which may be \emph{materialized} to local persistent storage). This simple subscription system provides a flexible framework for building complex event-driven application logic. In the remdainder of this section, we outline the specifics of the Textile Threads internals, focusing on three core components: Actions, the Dispatcher, and the Read/View Stores.

\subsubsection{Actions}

Like with Flux/Redux, to add an Event to the internal Threads system, we use \emph{Actions}. Actions cause predictable updates to state via \emph{Reducer} functions. In practice, Actions are \emph{dispatched} from a singleton \emph{Dispatcher}, which then stores the Event in an \emph{Event Store}, and calls a set of pure (in the functional sense) Reducer functions that mutate a set of downstream \emph{Read/View Stores}.

In addition to dispatching 'raw' Actions, application developers are able to build and interact with ction \emph{Creators}, which are simple functions (or groups of functions) that create (and possibly dispatch) new Actions. For those familiar with DDD, if built around a specific \emph{domain}, Action Creators provide bounded context that can be (roughly) compared to an aggregate root (i.e., a cluster of objects, that create a clear reference to the root aggregate). This is a powerful design feature that we will take advantage of in section \dots.

\subsubsection{Dispatcher}

In order to actually \emph{store} events and dispatch actions to downstream views, a Dispatcher is used. The dispatcher wraps an embedded Event Store, which is responsible for persisting Events. This means a Dispatcher is both an Event Store an Event Bus/Dispatcher. In practice, the internal Event Store supports ACID transactions [\dots], so that Actions are only ever dispatched if they were successfully persisted to the Event Store.

All ``side effects'' of a given Action happen within a `Transaction`, in order to ensure consistently of both storage (Event Store) and actions. Transactions provide a similar experience to Redux ``Sagas'' [\dots], but with a simpler interface. Once an Event has been persisted in the internal Event Store, the Dispatcher is responsible for running a `Store`'s `Reducer` callback. In order to make this possible, Stores must \emph{registered} with the Dispatcher.

@@

\subsubsection{Read Models}

@@

@@

Similarly, Threads defines a ViewStore, which are a special kind of ReadStore, akin to a Redux Store. In practice, they are smaller, read-only (immutable) models, that are built on-the-fly, and defined by a pure Reducer function. Threads use ViewStores to define the ACL for a given Thread.

Like in Flux, Stores (as they are outlined above) have several specific properties:

@@

A database built on Threads would map database operations into \texttt{Actions}. Here, the \emph{domain} is the databse. These \texttt{Actions} would then mutate a \texttt{ReadModel} via its corresponding \texttt{Reducer} function, as per above. The \texttt{Store} could be persisted or kept in-memory (as a \texttt{ViewModel}), and would be directly queried via corresponding database-style \texttt{Resolver} metods. Because Textile uses ES patterns, any delete-type operation would simply lead to an internal \emph{tombstone} \texttt{Action}, marking the database item for removal from the \texttt{ReadStore}.

@@

A key/value store built on Threads would ``map'' these database \texttt{Actions} into key/value operations, such as \texttt{Put}, \texttt{Get}, and \texttt{Del}. These \texttt{Actions} would then mutate a map-like \texttt{ReadModel}. To complete the database-like API, the \texttt{ActionCreator} and \texttt{ReadModel} methods outlined in Section \dots could be encapsulated in a database structure that satisfies a given interface (see for example, Figure \ref{ex:KVStore}).

>>>>

To make Threads as easy to adopt and use as possible, we've designed the internal event store to take advantage of ideas from several existing CQRS and ES systems, as well as ideas and designs from Flux \cite{facebookFluxInDepthOverview2019} and Redux\footnote{Flux/Redux build on concepts from CRQS and ES themselves, and Redux is arguably an implementation of the Flux application architecture.} \cite{reduxMotivation} and domain driven design \cite{evansDomaindrivenDesignTackling2004a} (DDD)\footnote{\url{https://dddcommunity.org}}. In this section, we outline the internal structure of a local Threads ``client'' at a relatively high-level, and outline how it can be used to build intuitive developer-facing APIs to make adopting and using Threads ``the right choice'' for a wide range of developers.

@@

For developers familiar with common application state management issues, tools such as Flux/Redux provide opinionated frameworks for making state mutations \emph{predictable} by imposing certain restrictions on how and when updates can happen. If you take a look at the architecture of a Flux (or Redux) application, you'll notice several core similaries to CQRS and ES systems (see \cite{facebookFluxInDepthOverview2019} for example). In both cases, the focus is on unidirectional data flows that build downstream views from atomic updates in the form of events (or actions).

We adopt a similar flow in Threads (see Figure \dots), such that views are updated by subscribing to specific events (or actions as in Flux/Redux), and applying a reducer function that is used to update the view state (which may be \emph{materialized} to local persistent storage). This simple subscription system provides a flexible framework for building complex event-driven application logic. In the following, we outline the specifics of the Textile Threads internals, focusing on three core components: Actions, the Dispatcher, and Read/View Stores.

\subsubsection{Actions}\label{sec:actions}

Like with Flux/Redux, to add an Event to the internal Threads system, we use \emph{Actions}. Actions cause predictable updates to state via \emph{Reducer} functions. In practice, Actions are \emph{dispatched} from a singleton \emph{Dispatcher} (Section \ref{sec:dispatcher}), which then stores the Event in an \emph{Event Store}, and calls a set of pure (in the functional sense) Reducer functions that mutate a set of downstream \emph{Read/View Stores}.

In addition to dispatching 'raw' Actions, application developers are able to build and interact with ction \emph{Creators}, which are simple functions (or groups of functions) that create (and possibly dispatch) new Actions. For those familiar with DDD, if built around a specific \emph{domain}, Action Creators provide bounded context that can be (roughly) compared to an aggregate root (i.e., a cluster of objects, that create a clear reference to the root aggregate). This is a powerful design feature that we will take advantage of in section \ref{sec:AccessControl}.

\subsubsection{Dispatcher}\label{sec:dispatcher}

In order to actually \emph{store} events and dispatch actions to downstream views, a Dispatcher is used. The dispatcher wraps an embedded Event Store, which is responsible for persisting Events. This means a Dispatcher is both an Event Store an Event Bus/Dispatcher. In practice, the internal Event Store supports ACID transactions \cite{haerderPrinciplesTransactionorientedDatabase1983}, so that Actions are only ever dispatched if they were successfully persisted to the Event Store.

All ``side effects'' of a given Action happen within a \emph{Transaction}, in order to ensure consistently of both storage (Event Store) and actions. Transactions provide a similar experience to Redux ``Sagas''\footnote{\url{https://redux-saga.js.org/}}, but with a simpler interface. Once an Event has been persisted in the internal Event Store, the Dispatcher is responsible for running a Store's Reducer callback. In order to make this possible, Stores must \emph{registered} with the Dispatcher.

@@

\subsubsection{Read Models}\label{sec:readModels}

@@

Like in Flux, Stores (as they are defined here) have several specific properties:

Threads also define \emph{ViewStores}, which are a special kind of ReadStore, akin to a Redux Store. In practice, they are smaller, read-only (immutable) models, that are built on-the-fly, and defined by a pure Reducer function. Textile Threads use ViewStores to define the ACL for a given Thread.

\paragraph{Reducers}

@@

\paragraph{Resolvers}

@@

A database built on Threads would map database operations into Actions. Here, the \emph{domain} is the databse. These Actions would then mutate a ReadModel via its corresponding Reducer function, as per above. The Store could be persisted or kept in-memory (as a ViewModel), and would be directly queried via corresponding database-style Resolver metods. Because Textile uses ES patterns, any delete-type operation would simply lead to an internal \emph{tombstone} Action, marking the database item for removal from the ReadStore.

@@

A key/value store built on Threads would ``map'' these database Actions into key/value operations, such as \texttt{Put}, \texttt{Get}, and \texttt{Del}. These Actions would then mutate a map-like ReadModel. To complete the database-like API, the Action Creator and ReadModel methods outlined in Section \ref{sec:actions} could be encapsulated in a database structure that satisfies a given interface (see for example, Figure \ref{ex:KVStore}).
